### 2017-01-15 document/authorlist.tex
Notes on bursts

\author{J.~Selsing\inst{1}
\and XSGRB collaboration
%\and D.~Malesani\inst{1}
%\and T.~Kr\"{u}hler\inst{1}
%\and J.~P.~U. Fynbo\inst{1}
%\and P.~Goldoni\inst{14}
%\and A.~de~Ugarte~Postigo\inst{11}
%\and J.~Japelj\inst{20}
%\and P.~D'Avanzo
%\and Z.~Cano
%\and S.~Covino\inst{10}
%\and V.~D'Elia\inst{712}
%\and H.~Flores
%\and O.~E.~Hartoog\inst{6}
%\and J.~Hjorth\inst{1}
%\and P.~Jakobsson\inst{5}
%\and A.~Levan
%\and A.~Melandri
%\and S.~Piranomonte \inst{7}
%\and R.~S\'anchez-Ram\'irez\inst{11}
%\and S.~Schulze\inst{1718}
%\and N.~R.~Tanvir\inst{19}
%\and C.~Th{\"o}ne
%\and S.~D.~Vergani\inst{78}
%\and P.~M.~Vreeswijk\inst{3}
%\and D.~J.~Watson\inst{1}
%\and K.~Wiersema\inst{19}
%\and D.~Xu\inst{1}
%\and L.~Christensen\inst{1}
%\and A.~De~Cia\inst{3}
%\and L.~Kaper\inst{6}
%\and L.~A.~Antonelli
%\and F.~Fiore
%\and A.~Gomboc
%\and P.~Groot
%\and F.~Hammer
%\and C.~Ledoux\inst{2}
%\and E.~Maiorano
%\and B.~Milvang-Jensen\inst{1}
%\and E.~Palazzi
%\and E.~Pian
%\and J.~Schaye
%\and G.~Tagliaferri\inst{7},
%\and R.~A.~M.~J.~Wijers\inst{6}
}

\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark %1
%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
%\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
%\and Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany
%\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland
%\and Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands
%\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
%\and GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France
%\and American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA
%\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy
%\and Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain
%\and ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy
%\and Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic
%\and APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et Leonie Duquet, 75205 Paris Cedex 13, France
%\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
%\and Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3, 20126, Milano, Italy
%\and Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astronom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile
%\and Millennium Center for Supernova Science
%\and Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK
%\and University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI
}