### 2017-10-26 document/tables/sample_properties.tex

\begin{table*}[!ht]

	\centering
	\begin{tabular}{cccc}
		\hline
		\hline\noalign{\smallskip}
		{} & {Full \textit{Swift} sample} & {Statistical sample} &  {Followed up bursts} \\
		\hline\noalign{\smallskip}
		N$_{BAT}$ & 915 & 71 & 85\\
		$\log(15-150$keV fluence)  & $-5.9_{-0.7}^{+0.6}$ &  $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.6}^{+0.7}$   \\
		N$_{XRT}$ & 850 & 70 & 84\\
		$\log(0.3-10$keV flux) & $-12.3_{-0.8}^{+0.7}$ &  $-12.3_{-1.1}^{+0.4}$ &  $-12.4_{-0.6}^{+0.8}$  \\
		N$_{HI}$ & 878 & 70 & 85\\
		$\log(N_H)$ & $21.4_{-0.5}^{+0.5}$ &  $-21.2_{-0.4}^{+0.5}$ &  $21.4_{-0.7}^{+0.7}$  \\

		\hline\noalign{\smallskip}

	\end{tabular}
	\caption{Population properties for the \textit{Swift sample} and the subset of bursts fulfulling the sample criteria. The population characteristics for the three samples exhibit a large degree of similarity, signifying the ability of our selection criteria to effectively capture the underlying population.  \label{tab:sample_properties}}
\end{table*}