### 2017-11-24 document/tables/sample_properties.tex
Doing number to-dos and update sample comparison

\begin{table*}[!ht]

	\centering
	\begin{tabular}{cccc}
		\hline
		\hline\noalign{\smallskip}
		{} & {Full \textit{Swift} sample} & {Statistical sample} &  {Followed up bursts} \\
		\hline\noalign{\smallskip}
		N$_{BAT}$ & 987 & 156 & 89\\
		$\log(15-150$keV fluence)  & $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.6}^{+0.7}$ &  $-5.9_{-0.7}^{+0.7}$   \\
		N$_{XRT}$ & 909 & 154 & 88\\
		$\log(0.3-10$keV flux) & $-12.3_{-0.8}^{+0.7}$ &  $-12.4_{-0.8}^{+0.7}$ &  $-12.4_{-0.7}^{+0.9}$  \\
		N$_{\mathrm{HI_x}}$ & 250 & 95 & 74\\
		$\log(\mathrm{N}_{\mathrm{HI_x}})$ & $21.7_{-0.9}^{+0.6}$ &  $21.5_{-3.4}^{+0.7}$ &  $21.6_{-4.5}^{+0.7}$  \\

		\hline\noalign{\smallskip}

	\end{tabular}
	\caption{Population properties for the \textit{Swift sample} and the subset of bursts fulfulling the sample criteria. The population characteristics for the three samples exhibit a large degree of similarity, signifying the ability of our selection criteria to effectively capture the underlying population.  \label{tab:sample_properties}}
\end{table*}