### 2018-02-05 document/authorlist.tex
Final draft sent around

\author{
J.~Selsing\inst{1}
\and D.~Malesani
\and P.~Goldoni
\and J.~P.~U. Fynbo
\and T.~Kr\"{u}hler
\and A.~Antonelli
\and M.~Arabsalmani
\and J.~Bolmer
\and Z.~Cano
\and L.~Christensen
\and S.~Covino
\and P.~D'Avanzo
\and V.~D'Elia
\and A.~De~Cia
\and A.~de~Ugarte~Postigo
\and H.~Flores
\and M.~Friis
\and A.~Gomboc
\and J.~Greiner
\and P.~Groot
\and F.~Hammer
\and O.E.~Hartoog
\and K.~E.~Heintz
\and J.~Hjorth
\and P.~Jakobsson
\and J.~Japelj
\and L.~Kaper
\and C.~Ledoux
\and G.~Leloudas
\and A.~Levan
\and E.~Maiorano
\and A.~Melandri
\and B.~Milvang-Jensen
\and E.~Palazzi
\and J.T.~Palmerio
\and D.~Perley
\and E.~Pian
\and S. ~Piranomonte
\and G.~Pugliese
\and R.~S\'{a}nchez-Ram\'{\i}rez
\and S.~Savaglio
\and P.~Schady
\and S.~Schulze
\and J.~Sollerman
\and M.~Sparre
\and G.~Tagliaferri
\and N.~R.~Tanvir
\and C.~C.~Th\"{o}ne
\and S.D.~Vergani
\and P.~Vreeswijk
\and D.~Watson
\and K.~Wiersema
\and R.~Wijers
\and D.~Xu
}

\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark %1
%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
%\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
%\and Th\"uringer Landessternwarte Tautenburg, Sternwarte 5, 07778 Tautenburg, Germany
%\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, IS-107 Reykjavik, Iceland
%\and Astronomical Institute Anton Pannekoek, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, the Netherlands
%\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
%\and GEPI-Observatoire de Paris, CNRS UMR 8111, Univ. Paris-Diderot, 5 Place Jules Janssen - 92190 Meudon, France
%\and American River College, Physics and Astronomy Dpt., 4700 College Oak Drive, Sacramento, CA 95841, USA
%\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate, Italy
%\and Instituto de Astrof\'{\i}sica de Andaluc\'{\i}a (IAA-CSIC), Glorieta de la Astronom\'{\i}a s/n, 18008, Granada, Spain
%\and ASI-Science Data Centre, Via Galileo Galilei, I-00044 Frascati, Italy
%\and Institute of Experimental and Applied Physics, Czech Technical University in Prague, Horska 3a/22, 128 00 Prague 2, Czech Republic
%\and APC, Astroparticules et Cosmologie, Universite Paris Diderot, CNRS/IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cite, 10, Rue Alice Domon et Leonie Duquet, 75205 Paris Cedex 13, France
%\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
%\and Universit\`a degli studi di Milano-Bicocca, Piazza della Scienza 3, 20126, Milano, Italy
%\and Pontificia Universidad Cat\'{o}lica de Chile, Departamento de Astronom\'{\i}a y Astrof\'{\i}sica, Casilla 306, Santiago 22, Chile
%\and Millennium Center for Supernova Science
%\and Department of Physics and Astronomy, University ofLeicester, University Road, Leicester, LE1 7RH, UK
%\and University of Ljubljana,Department of Physics,Faculty of Mathematics \& Physics,SI
}