### 2018-02-15 document/authorlist.tex

\author{
J.~Selsing \inst{1}$^,$\thanks{\email{jselsing@dark-cosmology.dk}}
\addtocounter{footnote}{2}
\and D.~Malesani \inst{1, 2, 3}\fnmsep\thanks{On-call observer}
\and P.~Goldoni \inst{4}\fnmsep$^\dagger$
\and J.~P.~U. Fynbo \inst{1, 2}\fnmsep$^\dagger$
\and T.~Kr\"{u}hler \inst{5}\fnmsep$^\dagger$
\and L.~A.~Antonelli \inst{6}\fnmsep$^\dagger$
\and M.~Arabsalmani \inst{7, 8}
\and J.~Bolmer \inst{5, 9}\fnmsep$^\dagger$
\and Z.~Cano \inst{10}\fnmsep$^\dagger$
\and L.~Christensen \inst{1}
\and S.~Covino \inst{11}\fnmsep$^\dagger$
\and P.~D'Avanzo \inst{11}\fnmsep$^\dagger$
\and V.~D'Elia \inst{12}\fnmsep$^\dagger$
\and A.~De~Cia \inst{13}
\and A.~de~Ugarte~Postigo \inst{1, 10}\fnmsep$^\dagger$
\and H.~Flores \inst{14}\fnmsep$^\dagger$
\and M.~Friis \inst{15, 16}
\and A.~Gomboc \inst{17}
\and J.~Greiner \inst{5}
\and P.~Groot \inst{18}
\and F.~Hammer \inst{14}
\and O.E.~Hartoog \inst{19}\fnmsep$^\dagger$
\and K.~E.~Heintz \inst{1, 2, 20}\fnmsep$^\dagger$
\and J.~Hjorth \inst{1}\fnmsep$^\dagger$
\and P.~Jakobsson \inst{20}\fnmsep$^\dagger$
\and J.~Japelj \inst{19}\fnmsep$^\dagger$
\and D.~A.~Kann \inst{10}\fnmsep$^\dagger$
\and L.~Kaper \inst{19}
\and C.~Ledoux \inst{9}
\and G.~Leloudas \inst{1}
\and A.J.~Levan \inst{21}\fnmsep$^\dagger$
\and E.~Maiorano \inst{22}
\and A.~Melandri \inst{11}\fnmsep$^\dagger$
\and B.~Milvang-Jensen \inst{1, 2}
\and E.~Palazzi \inst{22}
\and J.~T.~Palmerio \inst{23}\fnmsep$^\dagger$
\and D.~A.~Perley \inst{24}\fnmsep$^\dagger$
\and E.~Pian \inst{22}
\and S. ~Piranomonte \inst{6}\fnmsep$^\dagger$
\and G.~Pugliese \inst{19}\fnmsep$^\dagger$
\and R.~S\'{a}nchez-Ram\'{\i}rez \inst{25}\fnmsep$^\dagger$
\and S.~Savaglio \inst{26}
\and P.~Schady \inst{5}
\and S.~Schulze \inst{27}\fnmsep$^\dagger$
\and J.~Sollerman \inst{28}
\and M.~Sparre \inst{29}\fnmsep$^\dagger$
\and G.~Tagliaferri \inst{11}
\and N.~R.~Tanvir \inst{30}\fnmsep$^\dagger$
\and C.~C.~Th\"{o}ne \inst{10}
\and S.D.~Vergani \inst{14}\fnmsep$^\dagger$
\and P.~Vreeswijk \inst{18, 26}\fnmsep$^\dagger$
\and D.~Watson \inst{1, 2}\fnmsep$^\dagger$
\and K.~Wiersema \inst{21, 30}\fnmsep$^\dagger$
\and R.~Wijers \inst{19}
\and D.~Xu \inst{31}\fnmsep$^\dagger$
\and T.~Zafar \inst{32}
}

\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark %1
%\and Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
\and The Cosmic Dawn Center, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
\and DTU Space, National Space Institute, Technical University of Denmark, Elektrovej 327, 2800 Lyngby, Denmark
\and APC, Astroparticule et Cosmologie, Universit\'e Paris Diderot, CNRS/IN2P3, CEA/Irfu, Observatoire de Paris, Sorbonne Paris Cit\'e, 10, Rue Alice Domon et L\'eonie Duquet, 75205, Paris Cedex 13, France
\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
\and INAF - Osservatorio Astronomico di Roma, Via Frascati 33, 00078, Monte Porzio Catone (Roma), Italy
\and IRFU, CEA, Universit\'e Paris-Saclay, F-91191 Gif-sur-Yvette, France
\and Universit\'e Paris Diderot, AIM, Sorbonne Paris Cit\'e, CEA, CNRS, F-91191 Gif-sur-Yvette, France
\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain. % 10
%
\and INAF - Osservatorio Astronomico di Brera, via Bianchi 46, 23807, Merate (LC), Italy
\and Space Science Data Center - Agenzia Spaziale Italiana, via del Politecnico, s.n.c., I-00133, Roma, Italy
\and European Southern Observatory, Karl-Schwarzschild Str. 2, 85748 Garching bei M\”unchen, Germany
\and GEPI, Observatoire de Paris, PSL University, CNRS, 5 place Jules Janssen 92195, Meudon, France
\and KTH Royal Institute of Technology, Department of Physics, 106 91 Stockholm, Sweden
\and The Oskar Klein Centre for Cosmoparticle Physics, AlbaNova University Centre, 106 91 Stockholm, Sweden
\and Center for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 11c, 5270 Ajdov\v s\v cina, Slovenia
\and Department of Astrophysics, IMAPP, Radboud University Nijmegen, PO Box 9010, 6500 GL Nijmegen, the Netherlands
\and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, the Netherlands
\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107, Reykjavik, Iceland
\and Department of Physics, University of Warwick, Coventry, CV4 7AL, UK % 20
%
\and IASF/INAF Bologna, via Piero Gobetti 101, 40129 Bologna, Italy
\and Sorbonne Universit\'e, CNRS, UMR7095, Institut d'Astrophysique de Paris, F-75014, Paris, France
\and Astrophysics Research Institute, Liverpool John Moores University, IC2, Liverpool Science Park, 146 Brownlow Hill, Liverpool L3 5RF, UK
\and INAF, Istituto Astrofisica e Planetologia Spaziali, Via Fosso del Cavaliere 100, I-00133 Roma, Italy
\and Physics Department, University of Calabria, I-87036 Arcavacata di Rende, Italy
\and Department of Particle Physics and Astrophysics, Weizmann Institute of Science, 234 Herzl Street, Rehovot, 761000, Israel
\and Department of Astronomy, Stockholm University, AlbaNova, 10691 Stockholm, Sweden 
\and Institut f\"ur Physik und Astronomie, Universit\"at Potsdam, Karl-Liebknecht-Str.\,24/25, 14476 Golm, Germany
\and Department of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, United Kingdom
\and CAS Key Laboratory of Space Astronomy and Technology, National Astronomical Observatories, Chinese Academy of Sciences, Beijing, 100012, P.R. China % 30
%
\and Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.
}