### 2018-09-21 document/referee_response2.txt
Addressing referee comments

# 
# ------------------- 
# Referee report 
# 
# This report presents a review of the article "The X-shooter GRB 
# afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The 
# authors present and make available to the community a complete sample 
# of GRB afterglows observed with the X-Shooter spectrograph on the 
# VLT. This dataset is remarkable in several respect (>100 spectra 
# available taken over 8 years) and some of these have not be published 
# elsewhere. However, there are a number of important issues regarding 
# both the scientific measurements and their presentation which are 
# detailed below. These concerns need to be addressed before the paper 
# can be considered for publication. More details on these points 
# together with other comments are provided below. 
# 
# 
# 
# Major Concerns: 
# 
# - One aspect of these data that needs more emphasis is the 
# intervening absorbers. The data presented in this paper present a 
# remarkable number of metal and HI intervening absorbers as can be 
# judged from reading the appendix. These deserve much more attention, 
# including a table listing their redshifts, column density 
# measurements when available, metal lines detected, etc. A figure 
# with their redshift distribution would also be very welcome. In 
# addition, it would be interesting to compare the properties of this 
# population with the absorbers towards background quasars.

This paper is meant as a general introduction of the XSGRB legacy sample
and constitutes the first in a series of papers. While we certainly
agree that the issue of intervening absorbers is an interesting one, we
feel that it is best treated in a dedicated publication, which we are
already working on (de Ugarte Postigo et al. in prep.). This paper will
contain the complete list of intervening systems, as well as linelists,
redshifts, EWs, etc..

# - The manuscript states that the sample is homogeneous which is a bit 
# misleading. The spectrograph used to make the follow-up is the same 
# for all afterglows, but the SNR (due to a combination of brightness 
# of the source and time of observations after the burst) and the 
# rest-frame wavelength coverage vary enormously. While remarkable in 
# many respects, it is not appropriate to qualify this sample as 
# homogeneous.

The referee is correct here. Our statement about homogeneity concerns
only the selection, not the outcome of our observations. We have
attempted to reword anywhere relevant and we added an explicit caveat on
the resulting S/N homogeneity in paragraph §2.1.

# - There are a great many details provided about the data reduction, some of 
# which the reader fails to relate to the final scientific goals. For 
# example, a correction is applied to correct offsets in the 
# wavelength calibration, which presumably mostly impact the redshift 
# determination. How does the improvement translate into the redshift 
# determination accuracy? Similarly, for the telluric correction. What 
# is the impact of this on the final science? Conversely, a 
# description of the flux calibration error estimate would have been 
# welcome.

Concerning the wavelength calibration improvement, we have quantified
the magnitude of the improvement, and explained its relevance (§3.2). We
have also added a short paragraph to §3.5 explaining the impact of the
telluric correction on the inferred results.

Regarding the flux calibration accuracy, we have added a new section ($3.7). This includes

# - The comparison of N(HI) distributions between intervening quasar 
# absorbers and intrinsic GRB absorbers is not direct. The measurement 
# of the column density of a system in the case of a quasar absorber 
# integrate across the whole system. Conversely, in the case of GRBs, 
# the burst event takes place at a random position in the host galaxy 
# and in fact, on average, the column density will measure over only 
# half of the system. This geometric effect should be taken into 
# account.

This is certainly true, when comparing QSO DLAs to GRB DLA. To accounted for the geometric effect in Fig.

#
# - Another concern is to do with using absorption features to measure 
# the redshift. Indeed, one cannot exclude that the absorption is not 
# at the position of host galaxy but lying in front of it. For this 
# reason, the redshifts provided should be seen as upper limits. 
#

# 
# Other Concerns: 
# 
# - the links to the actual data seem to be inactive. It is important to 
# be able to review the material which is made available to the 
# community. 
# 
# - caption of Fig. 3 is cut-out 
# 
# - Fig. 5: please label the y-axis 
# 
# - Fig. 10: caption is cut-out again 
# GRB150403A: could the continuum be lower? 
# GRB150915A: is this fit significant? 
# 
# - appendix A: refer to the spectral PSF in the title? 
# 
# - appendix B: science-verified, Paranalization: 
# perhaps avoid this ESO jargon? 
# 
# typos: 
# - pt. 4 p. 3: Paranl 
# - p. 3: exampl 
# - p. 3: sorrounding 
# - p. 9: curios 
# - there are numerous typos in the appendix which I will not review 
# here. Please make sure to double check spelling there. 
# 
# I would be happy to review an updated version of the manuscript where 
# the above comments are taken into account. 
# 
# -----------------------