### 2018-10-12 document/referee_response.txt
Finishing referee response

# 
# ------------------- Referee report
# 
# This report presents a review of the article "The X-shooter GRB
# afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The
# authors present and make available to the community a complete sample
# of GRB afterglows observed with the X-Shooter spectrograph on the VLT.
# This dataset is remarkable in several respect (>100 spectra available
# taken over 8 years) and some of these have not be published elsewhere.
# However, there are a number of important issues regarding both the
# scientific measurements and their presentation which are detailed
# below. These concerns need to be addressed before the paper can be
# considered for publication. More details on these points together with
# other comments are provided below.
# 
# # 
# # # 
# Major Concerns:
# 
# - One aspect of these data that needs more emphasis is the intervening
# absorbers. The data presented in this paper present a remarkable
# number of metal and HI intervening absorbers as can be judged from
# reading the appendix. These deserve much more attention, including a
# table listing their redshifts, column density measurements when
# available, metal lines detected, etc. A figure with their redshift
# distribution would also be very welcome. In addition, it would be
# interesting to compare the properties of this population with the
# absorbers towards background quasars. 
#

This paper is meant as a general introduction of the XSGRB legacy sample
and constitutes the first in a series of papers. While we certainly
agree that the issue of intervening absorbers is an interesting one, we
feel that it is best treated in a dedicated publication, which we have
already published Christensen et al., A&A, 608A, 84C, 2017,
(arXiv:1709.01084) and are working on de Ugarte Postigo et al. in
prep.). This paper will contain the complete list of intervening
systems, as well as linelists, redshifts, EWs, etc..

# - The manuscript states that the sample is homogeneous which is a bit
# misleading. The spectrograph used to make the follow-up is the same
# for all afterglows, but the SNR (due to a combination of brightness of
# the source and time of observations after the burst) and the
# rest-frame wavelength coverage vary enormously. While remarkable in
# many respects, it is not appropriate to qualify this sample as
# homogeneous.

The referee is correct here. Our statement about homogeneity concerns
only the selection, not the outcome of our observations. We have
attempted to reword anywhere relevant and we added an explicit caveat on
the resulting S/N homogeneity in paragraph §2.1.

# - There are a great many details provided about the data reduction,
# some of which the reader fails to relate to the final scientific
# goals. For example, a correction is applied to correct offsets in the
# wavelength calibration, which presumably mostly impact the redshift
# determination. How does the improvement translate into the redshift
# determination accuracy? Similarly, for the telluric correction. What
# is the impact of this on the final science? Conversely, a description
# of the flux calibration error estimate would have been welcome.

Concerning the wavelength calibration improvement, we have quantified
the magnitude of the improvement, and explained its relevance (§3.2). We
have also added a short paragraph to §3.6 explaining the impact of the
telluric correction on the inferred results.

Regarding the flux calibration accuracy, we have added a new section
($3.5). This includes a new way to correct for slit losses and tries to
assess the flux accuracy of the spectra presented. The paper has
significantly benefited from this suggestion.

# - The comparison of N(HI) distributions between intervening quasar
# absorbers and intrinsic GRB absorbers is not direct. The measurement
# of the column density of a system in the case of a quasar absorber
# integrate across the whole system. Conversely, in the case of GRBs,
# the burst event takes place at a random position in the host galaxy
# and in fact, on average, the column density will measure over only
# half of the system. This geometric effect should be taken into
# account.

This is certainly true, when comparing QSO DLAs to GRB DLA. We have
updated the figure and text surrounding Fig. 10 to reflect this. For
Fig. 9, where we are comparing GRB DLAs to GRB DLA, we have retained the
original values for N_HI.

# - Another concern is to do with using absorption features to measure
# the redshift. Indeed, one cannot exclude that the absorption is not at
# the position of host galaxy but lying in front of it. For this reason,
# the redshifts provided should be seen as upper limits. 
# 
# 
# 
# 
# # 
# Other Concerns:
# 
# - the links to the actual data seem to be inactive. It is important to
# be able to review the material which is made available to the
# community.

Completely true. These are now active. Especially the ESO phase 3
material have taken some time to get through. Apologies.

# - caption of Fig. 3 is cut-out

This is now fixed.

# - Fig. 5: please label the y-axis

Done, thanks.

# - Fig. 10: caption is cut-out again

Fixed, thanks.

# GRB150403A: could the continuum be lower?

We have remeasured the N_HI for this burst. It does not change the
inferred N_HI value, but the continuum now better matches the
measurement.

# GRB150915A: is this fit significant?

We agree this could be misinterpreted. This bursts has Lyman alpha
emission in the trough, as indicated in Tab. 4. We have now marked all
the bursts affect by Lyman alpha emission.

# - appendix A: refer to the spectral PSF in the title?

This has now been reworded.

# - appendix B: science-verified, Paranalization: perhaps avoid this ESO
# jargon?

We have tried to weed out the use of these types of words.

# typos: - pt. 4 p. 3: Paranl - p. 3: exampl - p. 3: sorrounding - p. 9:
# curios - there are numerous typos in the appendix which I will not
# review here. Please make sure to double check spelling there.

Thanks for noting these. They have been fixed.

# I would be happy to review an updated version of the manuscript where
# the above comments are taken into account.
# 
# -----------------------